On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-26 1_Introduction.tex
Updates from Overleaf

This research focus on existing JavaScript idioms that a previous 
research work have proved to introduce misunderstanding in \clang.
JavaScript the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

In particular, our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
To achieve that, we investigate JavaScript idioms that
might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
JavaScript---that is, small
pieces of code that hinder developers' understanding of the source code of a program
for which there are functionally equivalent alternatives that are easier
to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider
the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
%for presentation purpose only). 
The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang
language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a cleaned version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic
capabilities and weaker type system. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Here we investigate a group of atom candidates
that stems from the Gopstein et al. study and detail additional JavaScript constructs
that might introduce misunderstanding---according to the opinion of JavaScript
developers.

@@

To understand whether an atom of confusion introduces misunderstandings,
we surveyed 140 JavaScript developers. In the survey, 
their task was to predict the output of short code
snippets, with half of them containing atoms of confusion. 
{\color{red}The number of correct answers for the
  \emph{cleaned version} of the code snippets for seven (out of nine)
  atom candidates is at least 15\% higher than the corresponding
  \emph{confuse version}
  of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
  might also affects the time necessary for submitting a correct answer.
  For instance, the presence of the Comma Operator atom increases
  the average time to predict the output by 65\%, though the output of snippets containing
  the Pre-Increment atom were correctly predicted 44\% faster.

  To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
  experienced professional developers. Participants considered the code snippets without
  the atom candidates easier to understand in 70\% of the cases,
  therefore corroborating the survey findings. The participants also discussed other 
  idioms that might lead developers to misunderstand JavaScript code.
  The occurrence of atoms in open-source repositories might reveal cases where developers
  should be more or less careful when writing code---and open the possibility to implement
  new program transformations for cleaning up confuse code. To identify such cases, we mined popular JavaScript projects
  from GitHub and found four atom candidates that are highly used in JavaScript
  code: the Implicit Predicate atom, for instance, occurs 19.89 times per KLOC of
  code.}

>>>>

%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.
JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

%In particular, 
Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer. To achieve that, we investigate JavaScript idioms that might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small
pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier
to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
%for presentation purpose only). 
The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang
language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a clean version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic
capabilities and weaker type system. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript
developers.

@@

To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers. In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion. 
{\color{red}The number of correct answers for the
  \emph{clean version} of the code snippets for 7 out of \na
  atom candidates is at least 15\% higher than the corresponding
  \emph{confusing version}
  of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
  might also affect the time necessary for submitting a correct answer.
  For instance, the presence of the Comma Operator atom increases
  the average time to predict the output by 76.23\%, though the outputs of snippets containing
  the Pre-Increment atom were correctly predicted 38.19\% faster.

  To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
  experienced professional developers. Participants considered the code snippets without
  the atom candidates easier to understand in 70\% of the cases,
  therefore corroborating the survey findings. The participants also discussed other 
  idioms that might lead developers to misunderstand JavaScript code.
  The occurrence of atoms in open-source repositories might reveal cases where developers
  should be more or less careful when writing code---and open the possibility to implement
  new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects
  from GitHub and found four atom candidates that are highly used in JavaScript
  code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.}

### 2021-04-26 2_Background.tex
Updates from Overleaf

As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs. Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different working settings. These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.

@@

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

@@

The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss between themselves the studied atoms.
The authors argue that this method is important to complement a quantitative analysis.
Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
}. \castor{We use interviews, but did we "test" the interviewees to verify if they understand code?} We follow a similar approach as we augment our quantitative analysis with interviews.

>>>>

As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs. Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different settings. These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.

@@

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

@@

The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms.
The authors argue that this method is important to complement a quantitative analysis.
Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
}. We follow a similar approach as we augment our quantitative analysis with interviews.

### 2021-04-26 3_Method.tex
Updates from Overleaf

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work. To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.

@@

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs. Our supplementary material
summarizes all atoms we consider in our research.

@@

The design of our first study blocks two variables
(developer experience and the code snippets) and
considers two treatments: the presence or absence of 
atoms of confusion within the code snippets. 
To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

@@

Having selected 10 atom candidates in our first study, we wrote 10 short programs, one containing each atom candidate. We call them the confusing versions of the programs. We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs. Overall, the survey employed 20 code snippets, 10 confusing and 10 clean. In order to reduce the cognitive effort, each subject was asked to predict the output of 10 code snippets, five confusing and five clean. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not).  We measured answer correctness and the total time each participant needed to answer the survey's questions.

@@

We implemented our survey as a web application. As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average. Fellow undergraduate students, professional colleagues, and friends took the pilot survey. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares.

@@

We decided not to provide feedback about the time students took to answer each question. Nor did we tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, if we gave respondents instant feedback, they might post comments on particular atoms, therefore interfering with future participants' thought process.

@@

We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
Its discussion threads, often called subreddits, are sorted by subject. For our research, we posted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey. Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares. We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets. Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey. When this happened, the corresponding row in the square was left incomplete. We considered all squares that contained incomplete rows to be invalid during data analysis and discarded them. Since we had\
 a large enough number of samples, the squares we had to discard did not impact our analysis.

@@

In the third part of the interview, participants were shown the code snippets that were used in the first study. The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand. The interviewers avoided introducing bias in the answers by not explaining that one of the sides of each pair contained the atom under investigation. Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.

@@

We first transcribed each interview and then examined the broad distribution of the answers. Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular. We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants. We present these results in Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.

@@

To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

>>>>

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions---though we postpone these results to future research work. To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.

@@

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs. Our supplementary material
summarizes all atoms we consider in our research.

@@

The design of our first study blocks two variables
(developer experience and the code snippets) and
considers two treatments: the presence or absence of 
atoms of candidates within the code snippets. 
To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of the code snippets are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

@@

Having selected 10 atom candidates in our first study, we wrote 10 short programs, one containing each atom candidate. We call them the confusing versions of the programs. We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs. Overall, the survey employed 20 code snippets, 10 confusing and 10 clean. In order to reduce the cognitive effort, each subject was asked to predict the output of 10 code snippets, five confusing and five clean. The order in which the questions were presented was randomized. By doing this, we seek to minimize the chances of subjects being aware that the current listing they are analyzing contains (or not) atoms of confusion. That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets do not).  We measured answer correctness and the total time each participant needed to answer the survey's questions.

@@

We implemented our survey as a web application. As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average. Undergraduate students and professional colleagues took the pilot survey. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares.

@@

We decided not to provide feedback about the time students took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, if we gave respondents instant feedback, they might post comments on particular atoms, therefore interfering with future participants' thought process.

@@

We posted the survey on Reddit.\footnote{Reddit is a North American online discussion platform.
Its discussion threads, often called subreddits, are sorted by subject. For our research, we posted information and links to the survey in two JavaScript subreddits.} We explained our research purposes and asked developers of any level of expertise to take the survey. Within twelve hours, we collected more than 150 answers, populating more than 70 replicas of the Latin Squares. We collected significant data on time taken and discrepancies in answer correctness between confusing and non-confusing versions of the snippets. Some inconsistencies arose while building the squares, for instance, when a user quit in the middle of the survey. When this happened, the corresponding row in the square was left incomplete. We considered all squares that contained incomplete rows to be invalid during data analysis and discarded them. 
%Since we had a large enough number of samples, the squares we had to discard did not impact our analysis.

@@

In the third part of the interview, participants were shown the code snippets that were used in the first study. The snippets were presented in pairs, where each pair contained a clean and a confusing version. The participants 
%were told both sides of each pair had the same output, and 
were asked to evaluate which version of the code was easier to understand. To avoid introducing bias in the answers, the interviewers did not explain that one of the versions in each pair contained the atom under investigation. Subjects were just presented the pairs of snippets and allowed to take the necessary time to decide on the most readable snippet.

@@

We first transcribed each interview and then examined the broad distribution of the answers. Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular. We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants. We present these results in Section~\ref{sec:interview-results}. 
%Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.

@@

To understand how often the analyzed atom candidates appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The sizes of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.
